### 2022-09-15 main/chapter_searches.tex
minor tweaks

\begin{figure}[tp]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_n.pdf}
\end{subfigure}
\caption[
A Poisson modelling example in the natural order.
]{%
A Poisson modelling example with $t(n) = \pm n$
test statistics, $b = 4.3$ and $s = 3.0$.
(left) \pvalues,
and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
background and signal-plus-background models.
The $n$ test statistic is Neyman-Pearson optimal because it is monotonically
(linearly) related to the likelihood ratio.
Each $\cls$ is a \pvalue\ ratio.
By reducing the observed data to an inequality, the $p$-value selectively
censors some of its information.
}
\label{fig:searches_sb_n}
\end{figure}

@@

Although the statement in Section~\ref{sec:searches_fisher} that \pvalue\
ratios are not ``not \emph{necessarily} misleading'',
Figure~\ref{fig:searches_sb_n} shows a bias from the fact that
no data can give a $\cls$ that favours the signal model;
$\log\cls$ is non-positive with the $-n$ test statistic, and indeed non-negative
for the opposite ordering $t(n) = n$.
The $\log$ likelihood ratio on the actual data does not feature this bias:
it favours background-only for deficits and background-plus-signal for
excesses and is indeed linear in $n$:
\begin{equation}
\log \frac{L(s + b)}{L(b)}
= -s + n\log\!\left(1 + \frac{s}{b}\right)
.
\end{equation}
Since the likelihood ratio is linear in $n$, both $n$ and $-n$ are optimal
test statistics by the criteria of the Neyman-Pearson lemma.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=0.5\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
\caption{%
Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
}
\end{subfigure}
\\[.5em]
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_t.pdf}
\caption{%
(left) \pvalues\ for Poisson data,
and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
background and signal-plus-background models.%
}
\end{subfigure}
\caption[
Analysis of a minimal Poisson modelling example with maximum likelihood ratio
test statistics
]{%
Analysis of a minimal Poisson modelling example with maximum likelihood ratio
test statistics $t(n) = \pm \Lambda(n)$,
where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$; $b = 4.3$ and $s = 3.0$.
This test statistic jumbles up the data into an that makes $n=11$ appear to be
weaker support for the signal than $n=9$.
Each $\cls$ is \pvalue\ ratio.
By reducing the observed data to an inequality, the $p$-value selectively
censors some of its information.
}
\label{fig:searches_sb_t}
\end{figure}

>>>>

\begin{figure}[tp]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_n.pdf}
\end{subfigure}
\caption[
A Poisson model in the natural order
]{%
A Poisson model in the natural order of $t(n) = \pm n$
test statistics, $b = 4.3$ and $s = 3.0$.
(left) $\log$ \pvalues,
and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
background and signal-plus-background models.
The $n$ test statistic is Neyman-Pearson optimal because it is monotonically
(linearly) related to the likelihood ratio.
Each $\cls$ is a \pvalue\ ratio.
By reducing the observed data to an inequality, the $p$-value selectively
censors some of its information.
}
\label{fig:searches_sb_n}
\end{figure}

@@

Although the statement in Section~\ref{sec:searches_fisher} that \pvalue\
ratios are not ``not \emph{necessarily} misleading'',
Figure~\ref{fig:searches_sb_n} shows a bias from the fact that
no data can give a $\cls$ that favours the signal model;
$\log\cls$ is non-positive with the $-n$ test statistic, and indeed non-negative
for the opposite ordering $t(n) = n$.
The $\log$ likelihood ratio on the actual data does not feature this bias:
it favours background-only for deficits and background-plus-signal for
excesses.
Since
\begin{equation}
\log \frac{L(s + b)}{L(b)}
= -s + n\log\!\left(1 + \frac{s}{b}\right)
\end{equation}
is linear in $n$, both $n$ and $-n$ are optimal in the Neyman-Pearson sense.

@@

\begin{figure}[tp]
\centering
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=0.5\textwidth]{figures/searches_cls_plots_with_pvals_order.pdf}
\caption{%
Ordering the data by a maximum likelihood ratio test statistic with $b=4.3$.
}
\end{subfigure}
\\[.5em]
\begin{subfigure}{\textwidth}
\centering
\includegraphics[width=\textwidth]{figures/searches_cls_plots_with_pvals_t.pdf}
\caption{%
(left) $\log$ \pvalues\ for Poisson data,
and (right) $\log$ likelihood ratios and $\log$ $\cls$ for
background and signal-plus-background models.%
}
\end{subfigure}
\caption[
Analysis of a minimal Poisson model with maximum likelihood ratio
test statistics
]{%
Analysis of a minimal Poisson model with maximum likelihood ratio
test statistics $t(n) = \pm \Lambda(n)$,
where $\Lambda(n) = L(b)/\mathrm{max}_x\,L(x)$; $b = 4.3$ and $s = 3.0$.
This test statistic jumbles up the data into an that makes $n=11$ appear to be
weaker support for the signal than $n=9$.
Each $\cls$ is \pvalue\ ratio.
By reducing the observed data to an inequality, the $p$-value selectively
censors some of its information.
}
\label{fig:searches_sb_t}
\end{figure}
